\chapter*{Acknowledgement}

        I would like to thank all the people that contributed, directly or indirectly, to create the rich and interesting environment in which I had the chance to work on this thesis. First of all, I would like to thank my supervisors Romain Laroche and Fabrice Lef\`evre for their help and their support. Romain provided valuable ideas and advice that helped a lot along the way, as well as a refreshing point of view whenever I felt stuck. Fabrice was also very encouraging and his experienced eye raised critical questions which always led to improved versions of my publications and my work in general.

        During these three years, I also had the chance to meet several other PhD studients with which I developed both professional and personal relationships. I would like to thank Layla El Asri for her important and helpful advice, Tatiana Ekeinhor-Komi for always maintaining the good mood and the fun, Emmanuel Ferreira for his help and for welcoming me during my stay at the LIA, Merwan Barlier for always sharing the novelties in the field of artificial intelligence as well as interesting conferences, Nicolas Carrara for the interesting technical discussions. Also, I want to thank Tianxiang Xia for the great job he did during his internship with me at Orange.

        It would not have been possible for me to complete this thesis without the help, the trust and the autonomy I was given both at Orange and the LIA. The NaDia team I was enrolled in at Orange were extremely supportive, very interested in my thesis topic and very helpful and reactive when I needed advice concerning the dialogue solution we are working with: Disserto. On the other hand, the researchers at the LIA were very welcoming and interesting to work with.

        Finally, I would like to thank my parents for their support during the course of this thesis.

        
